---
title: "Final Group Project: Updated Checkpoint Submission"
author: "Group 6 - Harsh, Nick, Rahul, Dhruv"
date: "`r Sys.Date()`"
output: pdf_document
header-includes:
  \usepackage{fvextra}
  \DefineVerbatimEnvironment{Highlighting}{Verbatim}{breaklines,commandchars=\\\{\}}
---

# Introduction

Background and Stuff

# Research question 

The following are our research questions (with scope of further improvement), with some of them complete and some of them slated for future work.
1. What is the relationship between PM2.5 and electricity generation in counties?
2. Of the emissions that come from electricity generation, what are the most significant in contributing to CO2e?
3. What does the introduction of the transportation sector to explore PM2.5 links look like?
4. Analyzing impact of electricity generation related emissions for counties (future work). This impact can be related to biodiversity, income etc. Will discuss with TA for clarity.

#Data Preparation and Cleaning, along with Exploratory Data Analysis
#PM2.5 Data
```{r}
library(dplyr)
library(knitr)
library(tidyverse)

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

#Store all PM2.5 Data for all counties

PM2.5_2016 <- read.csv("C:/Users/Harsh Vaibhav/Desktop/Data Science Data Sets/Final/PM2.5 Data/daily_88101_2016.csv")
PM2.5_2018 <- read.csv("C:/Users/Harsh Vaibhav/Desktop/Data Science Data Sets/Final/PM2.5 Data/daily_88101_2018.csv")
PM2.5_2019 <- read.csv("C:/Users/Harsh Vaibhav/Desktop/Data Science Data Sets/Final/PM2.5 Data/daily_88101_2019.csv")
PM2.5_2020 <- read.csv("C:/Users/Harsh Vaibhav/Desktop/Data Science Data Sets/Final/PM2.5 Data/daily_88101_2020.csv")
PM2.5_2021 <- read.csv("C:/Users/Harsh Vaibhav/Desktop/Data Science Data Sets/Final/PM2.5 Data/daily_88101_2021.csv")
PM2.5_2022 <- read.csv("C:/Users/Harsh Vaibhav/Desktop/Data Science Data Sets/Final/PM2.5 Data/daily_88101_2022.csv")

#Extract year from date format in the CSV (MM/DD/YYYY) and group by year

#2016
date_2016 <- PM2.5_2016$Date.Local
date_2016_asDate <- substr(date_2016, start = 0, stop = 4)
PM2.5_2016$Year <- as.integer(date_2016_asDate)

PM2.5_2016 <- PM2.5_2016 %>% 
  group_by(County.Name,Year, State.Name) %>%
  summarise(PM2.5_Avg = mean(Arithmetic.Mean))

#2018
date_2018 <- PM2.5_2018$Date.Local
date_2018_asDate <- substr(date_2018, start = 0, stop = 4)
PM2.5_2018$Year <- as.integer(date_2018_asDate)

PM2.5_2018 <- PM2.5_2018 %>% 
  group_by(County.Name,Year, State.Name) %>%
  summarise(PM2.5_Avg = mean(Arithmetic.Mean))

#2019
date_2019 <- PM2.5_2019$Date.Local
date_2019_asDate <- substr(date_2019, start = 0, stop = 4)
PM2.5_2019$Year <- as.integer(date_2019_asDate)

PM2.5_2019 <- PM2.5_2019 %>% 
  group_by(County.Name,Year, State.Name) %>%
  summarise(PM2.5_Avg = mean(Arithmetic.Mean))

#2020
date_2020 <- PM2.5_2020$Date.Local
date_2020_asDate <- substr(date_2020, start = 0, stop = 4)
PM2.5_2020$Year <- as.integer(date_2020_asDate)

PM2.5_2020 <- PM2.5_2020 %>% 
  group_by(County.Name,Year, State.Name) %>%
  summarise(PM2.5_Avg = mean(Arithmetic.Mean))

#2021
date_2021 <- PM2.5_2021$Date.Local
date_2021_asDate <- substr(date_2021, start = 0, stop = 4)
PM2.5_2021$Year <- as.integer(date_2021_asDate)

PM2.5_2021 <- PM2.5_2021 %>% 
  group_by(County.Name,Year, State.Name) %>%
  summarise(PM2.5_Avg = mean(Arithmetic.Mean))

#2022
date_2022 <- PM2.5_2022$Date.Local
date_2022_asDate <- substr(date_2022, start = 0, stop = 4)
PM2.5_2022$Year <- as.integer(date_2022_asDate)

PM2.5_2022 <- PM2.5_2022 %>% 
  group_by(County.Name,Year, State.Name) %>%
  summarise(PM2.5_Avg = mean(Arithmetic.Mean))

# Visualizing PM2.5 Trends
TotalAvg_PM2.5 <- data.frame(
  Year = c("2016", "2018", "2019", "2020", "2021", "2022"),
  PM2.5_Avg = c(
    mean(PM2.5_2016$PM2.5_Avg),
    mean(PM2.5_2018$PM2.5_Avg),
    mean(PM2.5_2019$PM2.5_Avg),
    mean(PM2.5_2020$PM2.5_Avg),
    mean(PM2.5_2021$PM2.5_Avg),
    mean(PM2.5_2022$PM2.5_Avg)
  )
)

ggplot(data = TotalAvg_PM2.5) +
  geom_line(aes(x = Year, y = PM2.5_Avg, group = 1), colour = "black") +
  labs(title = "PM 2.5 Trends (2016-2022)",
       x = "Year",
       y = "PM 2.5 Levels (ug/m3)") +
  theme_minimal()

```

#eGRID Data
``` {r}
df <- "C:/Users/Harsh Vaibhav/Desktop/Data Science Data Sets/Final/PM2.5 Data/egrid_total_data.xlsx"

egrid_data <- readxl::read_excel(df)

egrid_data <- egrid_data %>% rename(Year = `Data Year`)

energy_data <- egrid_data %>%
  group_by(`Plant county name`, Year) %>%
  summarise(
    `Plant annual coal net generation (MWh)` = sum(`Plant annual coal net generation (MWh)`, na.rm = TRUE),
    `Plant annual oil net generation (MWh)` = sum(`Plant annual oil net generation (MWh)`, na.rm = TRUE),
    `Plant annual gas net generation (MWh)` = sum(`Plant annual gas net generation (MWh)`, na.rm = TRUE),
    `Plant annual nuclear net generation (MWh)` = sum(`Plant annual nuclear net generation (MWh)`, na.rm = TRUE),
    `Plant annual hydro net generation (MWh)` = sum(`Plant annual hydro net generation (MWh)`, na.rm = TRUE),
    `Plant annual biomass net generation (MWh)` = sum(`Plant annual biomass net generation (MWh)`, na.rm = TRUE),
    `Plant annual wind net generation (MWh)` = sum(`Plant annual wind net generation (MWh)`, na.rm = TRUE),
    `Plant annual solar net generation (MWh)` = sum(`Plant annual solar net generation (MWh)`, na.rm = TRUE),
    `Plant annual geothermal net generation (MWh)` = sum(`Plant annual geothermal net generation (MWh)`, na.rm = TRUE),
    `Plant annual other fossil net generation (MWh)` = sum(`Plant annual other fossil net generation (MWh)`, na.rm = TRUE),
    `Plant annual other unknown/ purchased fuel net generation (MWh)` = sum(`Plant annual other unknown/ purchased fuel net generation (MWh)`, na.rm = TRUE),
    `Plant annual total nonrenewables net generation (MWh)` = sum(`Plant annual total nonrenewables net generation (MWh)`, na.rm = TRUE),
    `Plant annual total renewables net generation (MWh)` = sum(`Plant annual total renewables net generation (MWh)`, na.rm = TRUE)
  )

#Visualizing the Generation
summarized_data <- energy_data %>%
  group_by(Year) %>%
  summarise(
    `Plant annual coal net generation (MWh)` = sum(`Plant annual coal net generation (MWh)`, na.rm = TRUE),
    `Plant annual oil net generation (MWh)` = sum(`Plant annual oil net generation (MWh)`, na.rm = TRUE),
    `Plant annual gas net generation (MWh)` = sum(`Plant annual gas net generation (MWh)`, na.rm = TRUE),
    `Plant annual nuclear net generation (MWh)` = sum(`Plant annual nuclear net generation (MWh)`, na.rm = TRUE),
    `Plant annual hydro net generation (MWh)` = sum(`Plant annual hydro net generation (MWh)`, na.rm = TRUE),
    `Plant annual biomass net generation (MWh)` = sum(`Plant annual biomass net generation (MWh)`, na.rm = TRUE),
    `Plant annual wind net generation (MWh)` = sum(`Plant annual wind net generation (MWh)`, na.rm = TRUE),
    `Plant annual solar net generation (MWh)` = sum(`Plant annual solar net generation (MWh)`, na.rm = TRUE),
    `Plant annual geothermal net generation (MWh)` = sum(`Plant annual geothermal net generation (MWh)`, na.rm = TRUE),
    `Plant annual other fossil net generation (MWh)` = sum(`Plant annual other fossil net generation (MWh)`, na.rm = TRUE),
    `Plant annual other unknown/ purchased fuel net generation (MWh)` = sum(`Plant annual other unknown/ purchased fuel net generation (MWh)`, na.rm = TRUE),
    `Plant annual total nonrenewables net generation (MWh)` = sum(`Plant annual total nonrenewables net generation (MWh)`, na.rm = TRUE),
    `Plant annual total renewables net generation (MWh)` = sum(`Plant annual total renewables net generation (MWh)`, na.rm = TRUE)
  )

ggplot(summarized_data, aes(x = Year)) +
  geom_line(aes(y = `Plant annual coal net generation (MWh)`, color = "Coal")) +
  geom_line(aes(y = `Plant annual oil net generation (MWh)`, color = "Oil")) +
  geom_line(aes(y = `Plant annual gas net generation (MWh)`, color = "Gas")) +
  geom_line(aes(y = `Plant annual nuclear net generation (MWh)`, color = "Nuclear")) +
  geom_line(aes(y = `Plant annual hydro net generation (MWh)`, color = "Hydro")) +
  geom_line(aes(y = `Plant annual biomass net generation (MWh)`, color = "Biomass")) +
  geom_line(aes(y = `Plant annual wind net generation (MWh)`, color = "Wind")) +
  geom_line(aes(y = `Plant annual solar net generation (MWh)`, color = "Solar")) +
  geom_line(aes(y = `Plant annual geothermal net generation (MWh)`, color = "Geothermal")) +
  geom_line(aes(y = `Plant annual other fossil net generation (MWh)`, color = "Other Fossil")) +
  geom_line(aes(y = `Plant annual other unknown/ purchased fuel net generation (MWh)`, color = "Other Unknown/Purchased Fuel")) +
  geom_line(aes(y = `Plant annual total nonrenewables net generation (MWh)`, color = "Total Nonrenewables")) +
  geom_line(aes(y = `Plant annual total renewables net generation (MWh)`, color = "Total Renewables")) +
  labs(title = "Electricity Generation by Year and Generation Type",
       x = "Year",
       y = "Total Generation (MWh)",
       color = "Generation Type") +
  theme_minimal()

#Filter for counties that have data for all 6 years
energy_data <- energy_data %>%
  group_by(`Plant county name`) %>%
  filter(n_distinct(Year) == 6)

#We have different number of counties for each year in this data set, this is a problem. Lets find the common counties so it is easier to map trends for equal amount of time duration.

common_county_names <- Reduce(intersect, list(PM2.5_2016$County.Name, PM2.5_2018$County.Name, PM2.5_2019$County.Name, PM2.5_2020$County.Name, PM2.5_2021$County.Name, PM2.5_2022$County.Name))

Filtered_county_2016 <- PM2.5_2016 %>% filter(County.Name %in% common_county_names) 
Filtered_county_2018 <- PM2.5_2018 %>% filter(County.Name %in% common_county_names) 
Filtered_county_2019 <- PM2.5_2019 %>% filter(County.Name %in% common_county_names) 
Filtered_county_2020 <- PM2.5_2020 %>% filter(County.Name %in% common_county_names)
Filtered_county_2021 <- PM2.5_2021 %>% filter(County.Name %in% common_county_names)
Filtered_county_2022 <- PM2.5_2022 %>% filter(County.Name %in% common_county_names) 
Filtered_county_energy <- energy_data %>% filter(`Plant county name` %in% common_county_names)

Filtered_county_energy <- Filtered_county_energy %>%
  rename(County.Name = `Plant county name`)

combined_df <- bind_rows(Filtered_county_2016, Filtered_county_2018, Filtered_county_2019, Filtered_county_2020, Filtered_county_2021, Filtered_county_2022) %>%
  arrange(County.Name, Year)

County_PM2.5_Energy_Data <- merge(combined_df, Filtered_county_energy, by = c("County.Name", "Year"))
```

# Feature Selection
Before building models, we identify the most relevant features for predicting our target variable.

```{r}

#Feature selection is not applicable to the current stage the project is in.

```

# Model Building

#Electricity Model 1: Electricity and PM2.5
```{r}

PM2.5 <- County_PM2.5_Energy_Data$PM2.5_Avg
NonRenewableEnergy <- County_PM2.5_Energy_Data$`Plant annual total nonrenewables net generation (MWh)`
RenewableEnergy <- County_PM2.5_Energy_Data$`Plant annual total renewables net generation (MWh)`
CountyName <- County_PM2.5_Energy_Data$County.Name

Total_Energy <- lm(PM2.5 ~ NonRenewableEnergy + RenewableEnergy + factor(CountyName))
summary(Total_Energy)

ggplot()+
  geom_point(aes(x= NonRenewableEnergy, y= PM2.5))+
  geom_smooth(aes(x= NonRenewableEnergy, y= PM2.5))

ggplot()+
  geom_point(aes(x= RenewableEnergy, y= PM2.5))+
  geom_smooth(aes(x= RenewableEnergy, y= PM2.5))


PM2.5 <- County_PM2.5_Energy_Data$PM2.5_Avg
Coal <- County_PM2.5_Energy_Data$`Plant annual coal net generation (MWh)`
Oil <- County_PM2.5_Energy_Data$`Plant annual oil net generation (MWh)`
NGas <- County_PM2.5_Energy_Data$`Plant annual gas net generation (MWh)`
Nuclear <- County_PM2.5_Energy_Data$`Plant annual nuclear net generation (MWh)`
Hydro <- County_PM2.5_Energy_Data$`Plant annual biomass net generation (MWh)`
Wind <- County_PM2.5_Energy_Data$`Plant annual wind net generation (MWh)`
Biomass <- County_PM2.5_Energy_Data$`Plant annual biomass net generation (MWh)`
Solar <- County_PM2.5_Energy_Data$`Plant annual solar net generation (MWh)`
Geothermal <- County_PM2.5_Energy_Data$`Plant annual geothermal net generation (MWh)`
CountyName <- County_PM2.5_Energy_Data$County.Name

Source_Energy <- lm(PM2.5 ~ Coal + Oil + NGas + Nuclear + Hydro + Wind + Solar + Biomass + Geothermal + factor(CountyName))
summary(Source_Energy)
```
By running the linear regression of PM2.5 on type (renewable/non renewable) of generation, we get inconclusive results. The standard deviation of both the input variables, Non Renewable Energy and Renewable Energy, is too high for us to make anything of these results. This may very well be due to the restricted and limited number of years this regression is being carried out for.

#Transportation Link
```{r}

#Are we missing something? Literature tells us transportation is a very important contributor to PM2.5 emission. Let's narrow down to Pennsylvania's counties for now and check.

Transportation_2018<- read.csv("C:/Users/Harsh Vaibhav/Desktop/Data Science Data Sets/Final/Transporation Data/tabula-ReportofRegistration2018.csv")
Transportation_2019<- read.csv("C:/Users/Harsh Vaibhav/Desktop/Data Science Data Sets/Final/Transporation Data/tabula-ReportofRegistration2019.csv")
Transportation_2020<- read.csv("C:/Users/Harsh Vaibhav/Desktop/Data Science Data Sets/Final/Transporation Data/tabula-ReportofRegistration2020.csv")
Transportation_2021<- read.csv("C:/Users/Harsh Vaibhav/Desktop/Data Science Data Sets/Final/Transporation Data/tabula-ReportofRegistration2021.csv")
Transportation_2022<- read.csv("C:/Users/Harsh Vaibhav/Desktop/Data Science Data Sets/Final/Transporation Data/tabula-ReportofRegistration2022.csv")

combined_transportation_df <- bind_rows(Transportation_2018, Transportation_2019, Transportation_2020, Transportation_2021, Transportation_2022) %>%
  arrange(COUNTY, YEAR)

combined_transportation_df$COUNTY <- str_to_title(combined_transportation_df$COUNTY)

common_transportation_county_names <- Reduce(intersect, list(combined_transportation_df$COUNTY, combined_df$County.Name))

Filtered_transportation_df <- combined_transportation_df %>% filter(COUNTY %in% common_transportation_county_names)

Filtered_transportation_PM2.5_df <- combined_df %>% 
  filter(County.Name %in% common_transportation_county_names) %>%
  filter(State.Name %in% c("Pennsylvania", "Ohio", "Texas", "Iowa", "Georgia", "New York")) %>% 
  filter(Year != 2016)

County_PM2.5_Transportation_Data <- merge(Filtered_transportation_df, Filtered_transportation_PM2.5_df, 
                                   c("County.Name", "Year"), c("COUNTY", "YEAR"))

# Trend of Cars in Allegheny County

ggplot(County_PM2.5_Transportation_Data %>% 
         filter(COUNTY == "Allegheny"), 
       aes(x = YEAR)) +
  geom_line(aes(y = GAS/10^2, color = "GAS (hundreds)")) +
  geom_line(aes(y = DIESEL/10, color = "DIESEL (tens)")) +
  geom_line(aes(y = ELECTRIC, color = "ELECTRIC")) +
  geom_line(aes(y = PROPANE, color = "PROPANE")) +
  geom_line(aes(y = HYBRID, color = "HYBRID")) +
  geom_line(aes(y = OTHER, color = "OTHER")) +
  geom_line(aes(y = NATURAL.GAS, color = "NATURAL.GAS")) +
  geom_line(aes(y = COMP.GAS, color = "COMP.GAS")) +
  geom_line(aes(y = FLEX.FUEL, color = "FLEX.FUEL")) +
  labs(title = "Change in number of Vehicles in PA",
       x = "Year",
       y = "Number of Vehicles",
       color = "") +
  theme_minimal()
```


#Model 1: Linear Regression
```{r}
PM2.5_Avg <- County_PM2.5_Transportation_Data$PM2.5_Avg
Gas <- County_PM2.5_Transportation_Data$GAS
Diesel <- County_PM2.5_Transportation_Data$DIESEL
Electric <- County_PM2.5_Transportation_Data$ELECTRIC
Propane <- County_PM2.5_Transportation_Data$PROPANE
Hybrid <- County_PM2.5_Transportation_Data$HYBRID
NaturalGas <- County_PM2.5_Transportation_Data$NATURAL.GAS
CompGas <- County_PM2.5_Transportation_Data$COMP.GAS
FlexFuel <- County_PM2.5_Transportation_Data$FLEX.FUEL
Vehicles <- County_PM2.5_Transportation_Data$VEHICLES
PACounty <- County_PM2.5_Transportation_Data$COUNTY

# Fit the linear regression model
emission_significance_transport <- lm(PM2.5_Avg ~ Gas + Diesel + Electric + Propane + Hybrid + NaturalGas + CompGas + FlexFuel + factor(PACounty), data = County_PM2.5_Transportation_Data)
summary(emission_significance_transport)

emission_significance_transport1 <- lm(PM2.5_Avg ~ Diesel + FlexFuel + factor(PACounty))
summary(emission_significance_transport1)

emission_significance_transport2 <- lm(PM2.5_Avg ~ Diesel + factor(PACounty))
summary(emission_significance_transport2)

emission_significance_transport3 <- lm(PM2.5_Avg ~ FlexFuel + factor(PACounty))
summary(emission_significance_transport3)

# Set seed for reproducibility
set.seed(123)

# Generate random indices for splitting data into train and test sets
train_indices <- sample(1:nrow(County_PM2.5_Transportation_Data), 0.8 * nrow(County_PM2.5_Transportation_Data))

# Create train and test datasets
train_data <- County_PM2.5_Transportation_Data[train_indices, ]
test_data <- County_PM2.5_Transportation_Data[-train_indices, ]

# Case 1: Looking at all the vehicles 
# Train the model on the entire data-set
AllVehicles <- lm(PM2.5_Avg ~ VEHICLES + factor(COUNTY), data = train_data)
summary(AllVehicles)

# Predict PM2.5 levels for different number of cars in Allegheny County
predicted_PM2.5 <- predict(AllVehicles, newdata = data.frame(VEHICLES = c(855788, 855789), COUNTY = "Allegheny"))

# Calculate PM2.5 values for an addition of 1 car
difference <- predicted_PM2.5[2] - predicted_PM2.5[1]
print(difference)

# Calculate MSE
predicted_PM2.5_allegheny_all <- predict(AllVehicles, newdata = test_data)
mse_allegheny_all <- mean((predicted_PM2.5_allegheny_all - test_data$PM2.5_Avg)^2)
print(mse_allegheny_all)

# Case 2: Looking specifically at Gas, Diesel and Electric Vehicles
# Train the model on the entire data-set
GDE <- lm(PM2.5_Avg ~ GAS + DIESEL + ELECTRIC + factor(COUNTY), data = train_data)
summary(GDE)

# Predict PM2.5 levels for different number of cars in Allegheny County
predicted_PM2.5 <- predict(GDE, newdata = data.frame(GAS = c(811383, 811384), DIESEL = c(20091, 20092), ELECTRIC = c(4574,4575), COUNTY = "Allegheny"))

# Calculate PM2.5 values for an addition of 1 gas, 1 diesel and 1 electric car
difference <- predicted_PM2.5[2] - predicted_PM2.5[1]
print(difference)

# Calculate MSE
predicted_PM2.5_allegheny_GDE <- predict(GDE, newdata = test_data)
mse_allegheny_GDE <- mean((predicted_PM2.5_allegheny_GDE - test_data$PM2.5_Avg)^2, na.rm=TRUE)
print(mse_allegheny_GDE)

# Case 3: Looking specifically at Diesel Vehicles
# Train the model on the entire data-set
Diesel_Cars <- lm(PM2.5_Avg ~ DIESEL + factor(COUNTY), data = train_data)
summary(Diesel_Cars)

# Predict PM2.5 levels for different number of cars in Allegheny County
predicted_PM2.5 <- predict(Diesel_Cars, newdata = data.frame(DIESEL = c(20091, 20092), COUNTY = "Allegheny"))

# Calculate the difference between predicted PM2.5 values to get the PM2.5 for an addition of 1 diesel car
difference <- predicted_PM2.5[2] - predicted_PM2.5[1]
print(difference)

# Calculate MSE
predicted_PM2.5_allegheny_diesel <- predict(Diesel_Cars, newdata = test_data)
mse_allegheny_diesel <- mean((predicted_PM2.5_allegheny_diesel - test_data$PM2.5_Avg)^2)
print(mse_allegheny_diesel)



# Visualizing Gas vs. PM2.5_Avg with color based on year
ggplot(County_PM2.5_Transportation_Data) +
  geom_point(aes(x = log(Gas), y = PM2.5_Avg, color = as.factor(YEAR))) +
  geom_smooth(aes(x = log(Gas), y = PM2.5_Avg, color = as.factor(YEAR))) +
  labs(title = "Gas vs. PM2.5_Avg", x = "Log(Gas)", y = "PM2.5_Avg")

# Visualizing Diesel vs. PM2.5_Avg with color based on year
ggplot(County_PM2.5_Transportation_Data) +
  geom_point(aes(x = log(Diesel), y = PM2.5_Avg, color = as.factor(YEAR))) +
  geom_smooth(aes(x = log(Diesel), y = PM2.5_Avg, color = as.factor(YEAR))) +
  labs(title = "Diesel vs. PM2.5_Avg", x = "Log(Diesel)", y = "PM2.5_Avg")

# Visualizing Electric vs. PM2.5_Avg with color based on year
ggplot(County_PM2.5_Transportation_Data) +
  geom_point(aes(x = log(Electric), y = PM2.5_Avg, color = as.factor(YEAR))) +
  geom_smooth(aes(x = log(Electric), y = PM2.5_Avg, color = as.factor(YEAR))) +
  labs(title = "Electric vs. PM2.5_Avg", x = "Log(Electric)", y = "PM2.5_Avg")

# Visualizing Vehicles vs. PM2.5_Avg with color based on year
ggplot(County_PM2.5_Transportation_Data) +
  geom_point(aes(x = log(Vehicles), y = PM2.5_Avg, color = as.factor(YEAR))) +
  geom_smooth(aes(x = log(Vehicles), y = PM2.5_Avg, color = as.factor(YEAR))) +
  labs(title = "Vehicles vs. PM2.5_Avg", x = "Log(Vehicles)", y = "PM2.5_Avg")

```

#Model 2: Random Forest
```{r}
library(randomForest)

# Remove rows with missing values
County_PM2.5_Transportation_Data1 <- na.omit(County_PM2.5_Transportation_Data)

# Split data into training and testing sets [80-20]
set.seed(123)  # For reproducibility
train_indices <- sample(nrow(County_PM2.5_Transportation_Data1), 0.8 * nrow(County_PM2.5_Transportation_Data1))
train_data <- County_PM2.5_Transportation_Data1[train_indices, ]
train_data <- na.omit(train_data)
test_data <- County_PM2.5_Transportation_Data1[-train_indices, ]
test_data <- na.omit(test_data)

# Case 1: Looking at all the vehicles
rf_model_all_vehicles <- randomForest(train_data$PM2.5_Avg ~ train_data$VEHICLES + factor(train_data$COUNTY), data = train_data)
print(rf_model_all_vehicles)

# Predict PM2.5 levels on test data
predicted_PM2.5_all_vehicles <- predict(rf_model_all_vehicles, newdata = test_data)

# Evaluate model performance (e.g., mean squared error)
mse_all_vehicles <- mean((predicted_PM2.5_all_vehicles - test_data$PM2.5_Avg)^2)
print(paste("Mean Squared Error (All Vehicles):", mse_all_vehicles))

# Case 2: Looking specifically at Gas, Diesel, and Electric Vehicles
rf_model_gde <- randomForest(PM2.5_Avg ~ GAS + DIESEL + ELECTRIC + factor(COUNTY), data = train_data)
print(rf_model_gde)

# Predict PM2.5 levels on test data
predicted_PM2.5_gde <- predict(rf_model_gde, newdata = test_data)

# Evaluate model performance
mse_gde <- mean((predicted_PM2.5_gde - test_data$PM2.5_Avg)^2)
print(paste("Mean Squared Error (Gas, Diesel, Electric):", mse_gde))

# Case 3: Looking specifically at Diesel Vehicles
rf_model_diesel <- randomForest(PM2.5_Avg ~ DIESEL + factor(COUNTY), data = train_data)
print(rf_model_diesel)

# Predict PM2.5 levels on test data
predicted_PM2.5_diesel <- predict(rf_model_diesel, newdata = test_data)

# Evaluate model performance
mse_diesel <- mean((predicted_PM2.5_diesel - test_data$PM2.5_Avg)^2)
print(paste("Mean Squared Error (Diesel Vehicles):", mse_diesel))




```

#Model 3: Ridge
```{r}
# Install and load the glmnet package
library(glmnet)

#Trial_all
# Prepare the predictor matrix
x <- model.matrix(~. -1, data = allegheny_data %>% select(GAS, DIESEL, ELECTRIC, PROPANE, HYBRID, OTHER, NATURAL.GAS, COMP.GAS, FLEX.FUEL))
y <- allegheny_data$PM2.5_Avg

# Perform cross-validated ridge regression
ridge_model_all <- cv.glmnet(x, y, alpha = 0)
best_lambda_all <- ridge_model_all$lambda.min
final_model_all <- glmnet(x, y, alpha = 0, lambda = best_lambda_all)

# Extract coefficients
coefficients_all <- coef(final_model_all)

# Predict PM2.5 levels
predicted_PM2.5_all <- predict(final_model_all, newx = x)

# Calculate mean squared error
mse_all <- mean((predicted_PM2.5_all - y)^2)

# Print results
print(coefficients_all)
print(mse_all)

# Case 1: Looking at all the vehicles
# Prepare the predictor matrix
x <- model.matrix(~. -1, data = allegheny_data %>% select(VEHICLES))
y <- allegheny_data$PM2.5_Avg

# Perform cross-validated ridge regression
ridge_model_all_vehicles <- cv.glmnet(x, y, alpha = 0)
best_lambda_all_vehicles <- ridge_model_all_vehicles$lambda.min
final_model_all_vehicles <- glmnet(x, y, alpha = 0, lambda = best_lambda_all_vehicles)

# Extract coefficients
coefficients_all_vehicles <- coef(final_model_all_vehicles)

# Predict PM2.5 levels
predicted_PM2.5_all_vehicles <- predict(final_model_all_vehicles, newx = x)

# Calculate mean squared error
mse_all_vehicles <- mean((predicted_PM2.5_all_vehicles - y)^2)

# Print results
print(coefficients_all_vehicles)
print(mse_all_vehicles)

# Case 2: Looking specifically at Gas, Diesel, and Electric Vehicles
# Prepare the predictor matrix
x <- model.matrix(~. -1, data = allegheny_data %>% select(GAS, DIESEL, ELECTRIC))
y <- allegheny_data$PM2.5_Avg

# Perform cross-validated ridge regression
ridge_model_gde <- cv.glmnet(x, y, alpha = 0)
best_lambda_gde <- ridge_model_gde$lambda.min
final_model_gde <- glmnet(x, y, alpha = 0, lambda = best_lambda_gde)

# Extract coefficients
coefficients_gde <- coef(final_model_gde)

# Predict PM2.5 levels
predicted_PM2.5_gde <- predict(final_model_gde, newx = x)

# Calculate mean squared error
mse_gde <- mean((predicted_PM2.5_gde - y)^2)

# Print results
print(coefficients_gde)
print(mse_gde)
summary(final_model_gde)

# Case 3: Looking specifically at Diesel Vehicles
# Prepare the predictor matrix
x <- model.matrix(~. -1, data = allegheny_data %>% select(DIESEL))
y <- allegheny_data$PM2.5_Avg

# Perform cross-validated ridge regression
ridge_model_diesel <- cv.glmnet(x, y, alpha = 0)
best_lambda_diesel <- ridge_model_diesel$lambda.min
final_model_diesel <- glmnet(x, y, alpha = 0, lambda = best_lambda_diesel)

# Extract coefficients
coefficients_diesel <- coef(final_model_diesel)

# Predict PM2.5 levels
predicted_PM2.5_diesel <- predict(final_model_diesel, newx = x)

# Calculate mean squared error
mse_diesel <- mean((predicted_PM2.5_diesel - y)^2)

# Print results
print(coefficients_diesel)
print(mse_diesel)

```
#Model 4: Lasso
```{r}
#Trial_all
# Prepare the predictor matrix
x <- model.matrix(~. -1, data = allegheny_data %>% select(GAS, DIESEL, ELECTRIC, PROPANE, HYBRID, OTHER, NATURAL.GAS, COMP.GAS, FLEX.FUEL))
y <- allegheny_data$PM2.5_Avg

# Perform cross-validated lasso regression
lasso_model_all <- cv.glmnet(x, y, alpha = 1)
best_lambda_all <- lasso_model_all$lambda.min
final_model_all <- glmnet(x, y, alpha = 1, lambda = best_lambda_all)

# Extract coefficients
coefficients_all <- coef(final_model_all)

# Predict PM2.5 levels
predicted_PM2.5_all <- predict(final_model_all, newx = x)

# Calculate mean squared error
mse_all <- mean((predicted_PM2.5_all - y)^2)

# Print results
print(coefficients_all)
print(mse_all)

#Just GDE
# Prepare the predictor matrix
x <- model.matrix(~. -1, data = allegheny_data %>% select(GAS, DIESEL, ELECTRIC))
y <- allegheny_data$PM2.5_Avg

# Perform cross-validated lasso regression
lasso_model_all <- cv.glmnet(x, y, alpha = 1)
best_lambda_all <- lasso_model_all$lambda.min
final_model_all <- glmnet(x, y, alpha = 1, lambda = best_lambda_all)

# Extract coefficients
coefficients_all <- coef(final_model_all)

# Predict PM2.5 levels
predicted_PM2.5_all <- predict(final_model_all, newx = x)

# Calculate mean squared error
mse_all <- mean((predicted_PM2.5_all - y)^2)

# Print results
print(coefficients_all)
print(mse_all)
```


# Analysis (interpretation)



# Conclusions and Future Work



